On the presentation of orthogonal groups over polynomial rings

### 2019-06-06 dl-paper.tex
More material

@@

\subsection{Lemma33}
Throughout this section $\Phi$ is a simply-laced irreducible root system and $A$ is a local ring with maximal ideal $M$.

For a simply-laced root system $\Phi$ and a commutative ring $R$, the Steinberg group $\St(\Phi,\,R)$ is generated by the formal symbols $x_{\alpha}(a)$ for $\alpha\in\Phi$, $a\in R$ subject to the Steinberg relations
\begin{align}
&x_{\alpha}(a)x_{\alpha}(b)=x_{\alpha}(a+b), \tag{R1}\\
&[x_{\alpha}(a),\,x_{\beta}(b)]=x_{\alpha+\beta}(N_{\alpha\beta}\,ab),\text{ for }\alpha+\beta\in\Phi, \tag{R2} \\
&[x_{\alpha}(a),\,x_{\beta}(b)]=1,\text{ for }\alpha+\beta\not\in\Phi\cup0. \tag{R3}
\end{align}

Let $I\trianglelefteq A$ be a (not neccesarily proper) ideal of a commutative ring $A$. Denote $R=\oplus_{d\in\mathbb Z}R_d\subseteq A[t,t\inv]$, where $R_d=It^d$, $d<0$, and $R_d=At^d$, $d\geq0$.

For such a ring $R$ one can assume in the definition of the Steinberg group that the generators $x_\alpha(a)$ and relations R1--R3 are defined only for homogeneous elements.

Consider the root system $\rD_l$ for $l\geq5$. It can be naturally embedded into the root system of type $\rC_l$ as a set of short roots.

Define truncated Steinberg groups $G_n$ for $n\geq1$ by the set of generators $x_\alpha(a)$, $\alpha\in\rD_l$, $a\in R_d$, $d\leq n$, and relations

\begin{align}
&\,\,x_{\alpha}(a)x_{\alpha}(b) =  x_{\alpha}(a+b),& a,b\in R_d,\ d\leq n, \tag{R$1_n$} \\
&\,[x_{\alpha}(a),\,x_{\beta}(b)]= x_{\alpha+\beta}(N_{\alpha\beta}\,ab),& a\in R_d,\ b\in R_e,\ d,e,d+e\leq n, \tag{R$2_n$} \\
&[x_{\alpha}(a),\,x_{\beta}(b)]= 1,& \alpha+\beta\not\in\rC_l\cup0,\ a\in R_d,\ b\in R_e,\ d,e\leq n, \tag{R$3.1_n$} \\
&[x_{\alpha}(a),\,x_{\beta}(b)]= 1,& \alpha+\beta\in\rC_l\setminus\rD_l,\ a\in R_d,\ b\in R_e,\ d,e,d+e\leq n. \tag{R$3.2_n$}
\end{align}

\begin{rem}
We can take $\mathbb R^l$ with a base $\chi_1,\ldots,\chi_l$ and visualise root systems
\[\rD_l=\{\pm\chi_i\pm\chi_j\mid1\leq i\neq j\leq l\}\]
and $\rC_l=\rD_l\cup\{\pm2\chi_i\mid1\leq i\leq l\}$. Then we immediately see that $\alpha+\beta\in\rC_l\setminus\rD_l$ exactly means that either $\{\alpha,\beta\}$ or $\{-\alpha,-\beta\}$ equals $\{\chi_i-\chi_j,\chi_i+\chi_j\}$. Then for some $\sigma$ from the Weyl group of $\rD_l$ we get 
\[ \{\sigma\alpha,\sigma\beta\}=\{\chi_{l-1}-\chi_l, \chi_{l-1}+\chi_l\}=\{\alpha_{l-1},\alpha_l\} \]
(the horns of the Dynkin diagram $\rD_l$). Another condition on $\alpha$ and $\beta$ equivalent to $\alpha+\beta\in\rC_l\setminus\rD_l$ is that they do not lie in any subsystem of $\rD_l$ of type $\rA_m$ for $m\geq4$.
\end{rem}

@@

\begin{comment}
\begin{corollary}
  $T_\alpha$ satisfy the relations $\text{R3}_{1,1}$ in the remaining case when $\alpha$ is perpendicular to $\beta$.
\end{corollary}

\begin{align*} {}^{x_\gamma(X^{-1})}x_\alpha(bX^{-1}) = {}^{x_\gamma(X^{-1})}[x_{-\beta}(b'X^{-1}), x_{\alpha+\beta}(1)] & \text{ by R$2_1$}\\
 = [x_{-\beta}(b'X^{-1}), {}^{x_\gamma(X^{-1})}x_{\alpha+\beta}(1)] & \text{ by R$3_1$ for sharp angles}\\
 = [[x_{-\beta-\gamma}(b''), x_\gamma(X^{-1})], {}^{x_\gamma(X^{-1})}x_{\alpha+\beta}(1)] & \text{ by R$2_1$}\\
 = [x_{-\beta-\gamma}(b''), [x_\gamma(X^{-1}), x_{\alpha+\beta}(1)]] & \text{ by R$3_0$}\\
 = x_{\alpha}(b'''X^{-1}) & \text{ by R$2_1$.} \end{align*}

>>>>

In this paper we will be mostly interested in the case when the Dynkin diagram of $\Phi$ is simply-laced, in which case the above relations further simplify to
\begin{align}
&x_{\alpha}(a)x_{\alpha}(b)=x_{\alpha}(a+b), \tag{R1}\\
&[x_{\alpha}(a),\,x_{\beta}(b)]=x_{\alpha+\beta}(N_{\alpha\beta}\,ab),\text{ for }\alpha+\beta\in\Phi, \tag{R2} \\
&[x_{\alpha}(a),\,x_{\beta}(b)]=1,\text{ for }\alpha+\beta\not\in\Phi\cup0. \tag{R3}
\end{align}

@@

\subsection{Lemma on the presentation of Steinberg groups}
%Throughout this section $\Phi$ is a root system of type $\rA_{\geq 4}$, $\rD_{\geq 5}$, $\rE_{6,7,8}$ and $A$ is a local ring with maximal ideal $M$.
Let $I\trianglelefteq A$ be an ideal of a commutative ring $A$.
We consider $A[X, X\inv]$ as a graded ring in which $X^{-1}$ has degree $1$.
We denote by $B = B(A, I)$ the subring $A[X^{-1}] + I[X] \subseteq A[X, X\inv]$ with the induced grading.
As an $A$-module $B$ decomposes as $\oplus_{d\in\mathbb Z}B_d$ where $B_d=I \cdot X^d$ for $d<0$, and $B_d=A \cdot X^d$ for $d\geq0$.

We need to introduce notation for certain relations of $\St(\Phi, B)$.
\begin{align}
&\,\,x_{\alpha}(a)x_{\alpha}(b) =  x_{\alpha}(a+b),& a,b\in B_d, \tag{R$1_d$} \\
&\,[x_{\alpha}(a),\,x_{\beta}(b)]= x_{\alpha+\beta}(N_{\alpha\beta} \cdot ab),& a\in B_d,\ b\in B_e, \tag{R$2_{d,e}$} \\
&[x_{\alpha}(a),\,x_{\gamma}(b)]= 1,& \alpha \perp \beta,\ a\in B_d,\ b\in B_e, \tag{R$3^\bot_{d,e}$} \\
&[x_{\alpha}(a),\,x_{\beta}(b)]= 1,& \alpha - \beta \in \Phi,\ a\in B_d,\ b\in B_e. \tag{R$3^\angle_{d,e}$}
\end{align}
%For such a ring $R$ one can assume in the definition of the Steinberg group that the generators $x_\alpha(a)$ and relations R1--R3 are defined only for homogeneous elements.

For $n\geq 1$ we define ``truncated'' Steinberg groups $\St^{\leq n}(\Phi, B)$ by the set of generators $x_\alpha(a)$, $\alpha\in \Phi$, $a\in B_d$, $d\leq n$, and relations 
\begin{itemize}
 \item $\text{R}1_{d}$ for $d\leq n$;
 \item $\text{R2}_{d,e}$ for $d,e \leq n$ such that $d+e \leq n$;
 \item $\text{R3}^\angle_{d, e}$ for $d, e\leq n$;
 \item $\text{R3}^\bot_{d, 0}$ for $d \leq n$.
 \end{itemize}

\begin{comment}
\begin{rem}
We can take $\mathbb R^l$ with a base $\chi_1,\ldots,\chi_l$ and visualise root systems
\[\rD_l=\{\pm\chi_i\pm\chi_j\mid1\leq i\neq j\leq l\}\]
and $\rC_l=\rD_l\cup\{\pm2\chi_i\mid1\leq i\leq l\}$. Then we immediately see that $\alpha+\beta\in\rC_l\setminus\rD_l$ exactly means that either $\{\alpha,\beta\}$ or $\{-\alpha,-\beta\}$ equals $\{\chi_i-\chi_j,\chi_i+\chi_j\}$. Then for some $\sigma$ from the Weyl group of $\rD_l$ we get 
\[ \{\sigma\alpha,\sigma\beta\}=\{\chi_{l-1}-\chi_l, \chi_{l-1}+\chi_l\}=\{\alpha_{l-1},\alpha_l\} \]
(the horns of the Dynkin diagram $\rD_l$). Another condition on $\alpha$ and $\beta$ equivalent to $\alpha+\beta\in\rC_l\setminus\rD_l$ is that they do not lie in any subsystem of $\rD_l$ of type $\rA_m$ for $m\geq4$.
\end{rem}

Everywhere below we set $t = X^{-1}$.
\begin{lemma} \label{claim2} In $\St^{\leq n}(\Phi, B)$ relations $\text{R3}^\angle_{d, e}$ imply $\text{R3}^{\bot,0}_{d, e}$ for $1\leq d\leq e \leq n$. \end{lemma}
\begin{proof}
Without loss of generality we may assume that $\alpha, \gamma$ fit into some $\mathsf{A}_3$-triple $(\alpha, \beta, \gamma)$.
 \begin{align*}
 {}^{x_\gamma(t^d)}x_\alpha(bt^{e}) = {}^{x_\gamma(t^d)}[x_{-\beta}(b't^{e}), x_{\alpha+\beta}(1)] & \text{ by $\text{R2}_{e,0}$}\\
 = [x_{-\beta}(b't^{e}), {}^{x_\gamma(t^d)}x_{\alpha+\beta}(1)] & \text{ by $\text{R3}^\angle_{d, e}$}\\
 = [[x_{-\beta-\gamma}(b''t^{e-d}), x_\gamma(t^d)], {}^{x_\gamma(t^d)}x_{\alpha+\beta}(1)] & \text{ by $\text{R2}_{e-d,d}$}\\
 = [x_{-\beta-\gamma}(b''t^{e-d)}), [x_\gamma(t^d), x_{\alpha+\beta}(1)]] & \text{ by~\eqref{eq:H1iii} and $\text{R3}^\bot_{e-d,0}$}\\
 = x_{\alpha}(b'''t^{e}) & \text{ by $\text{R2}_{0,d}$ and $\text{R2}_{e-d,d}$.} \end{align*}
Usual identities for structure constants imply (cf.~\cite[p.~12]{Re75}):
 $b'''=N_{-\beta-\gamma, \alpha+\beta+\gamma} \cdot N_{\gamma, \alpha+\beta} \cdot N_{-\beta-\gamma, \gamma} \cdot N_{-\beta, \alpha+\beta} \cdot b = b,$ from which the claim follows.
\end{proof}

\begin{lemma} \label{claim3}
 In $\St^{\leq n}(\Phi, B)$ relations $\bot_0(d, e)$, $\angle(d, d)$ and $\angle(e, e)$ imply $\bot(d, e)$ for $d\leq e \leq n$.
\end{lemma}
\begin{proof}
 \begin{align*} 
 x_{\alpha+\beta+\gamma}(abt^{e}) = {}^{x_{-\beta}(t^d)}\!x_{\alpha+\beta+\gamma}(abt^{e}) & \text{ by $\text{R3}^{\bot, 0}_{d, e}$}\\
 = {}^{x_{-\beta}(t^d)}\![x_{\alpha+\beta}(b't^{e-d}), x_\gamma(at^d)] & \text{ by $\text{R2}_{e-d,d}$} \\
 = [x_{\alpha}(b't^{e}) \cdot x_{\alpha+\beta}(b't^{e-d}), x_\gamma(at^d)] & \text{ by $\text{R2}_{e-d,d}$ and $\text{R3}^\angle_{d, d}$ }\\
 = {}^{x_{\alpha}(b't^{e})}\!x_{\alpha+\beta+\gamma}(abt^{e}) [x_{\alpha}(b't^{e}), x_\gamma(at^d)] & \text{ by~\eqref{eq:H1ii} and $\text{R2}_{e-d,d}$} \\ = x_{\alpha+\beta+\gamma}(abt^{e}) [x_{\alpha}(b't^{e}), x_\gamma(at^d)] & \text{ by $\text{R3}^\angle_{e, e}$. \qedhere}
\end{align*}
\end{proof}
\end{comment}

\begin{align}
 \label{eq:H1ii}  [ab, c] = {}^a[b, c] \cdot [a,c];&\\ %= [a,[b,c]]\cdot [b,c] \cdot [a,c];&\\
 \label{eq:H1iii} [a,c]   = 1    \text{ implies } [a, [b,c]] = [[a,b],{}^bc].&
\end{align}

\begin{align*} 
 [x_\alpha(\xi), x_{\gamma}(bt^{e})] = [x_\alpha(\xi), [x_{\beta+\gamma}(b't^{e-1}), x_{-\beta}(t)]] & \text{ by~$\text{R2}_{e-1,1}$} \\
 = [[x_\alpha(\xi), x_{\beta+\gamma}(b't^{e-1})], {}^{x_{\beta+\gamma}(b't^{e-1})}\!x_{-\beta}(t)] & \text{ by \eqref{eq:H1iii} and $\text{R3}^\angle_{0,1}$} \\
 = {}^{x_{\beta+\gamma}(b't^{e-1})}\![x_{\alpha+\beta+\gamma}(b't^{e-1}\xi), x_{-\beta}(t)] & \text{ by $\text{R2}_{0,e-1}$ and $\text{R3}^\angle_{e-1,e-1}$} \\
 = 1 & \text{ by $\text{R3}^\bot_{e-1,1}$.}
\end{align*}

@@

\begin{comment}
\begin{align*} {}^{x_\gamma(X^{-1})}x_\alpha(bX^{-1}) = {}^{x_\gamma(X^{-1})}[x_{-\beta}(b'X^{-1}), x_{\alpha+\beta}(1)] & \text{ by R$2_1$}\\
 = [x_{-\beta}(b'X^{-1}), {}^{x_\gamma(X^{-1})}x_{\alpha+\beta}(1)] & \text{ by R$3_1$ for sharp angles}\\
 = [[x_{-\beta-\gamma}(b''), x_\gamma(X^{-1})], {}^{x_\gamma(X^{-1})}x_{\alpha+\beta}(1)] & \text{ by R$2_1$}\\
 = [x_{-\beta-\gamma}(b''), [x_\gamma(X^{-1}), x_{\alpha+\beta}(1)]] & \text{ by R$3_0$}\\
 = x_{\alpha}(b'''X^{-1}) & \text{ by R$2_1$.} \end{align*}
